---
title: "Clocks (Coldplay) - Chroma vs Timbre"
format:
  html:
    theme: flatly
    toc: true
    toc-location: right
execute:
  echo: false
  warning: false
  message: false
---

## Introduction

In this section, I compare chroma and timbre representations for *Clocks* by Coldplay.  
This song is a useful example because it repeats the same piano pattern and chord loop for most of the track, while the arrangement and texture still change over time. That makes it a good case for comparing harmonic similarity and timbral similarity.

```{r}
library(tidyverse)
library(compmus)

chroma_raw <- read_csv("songs_csv/clocks_chroma.csv")
mfcc_raw <- read_csv("songs_csv/clocks_MFCC.csv")

dir.create("plots", showWarnings = FALSE)
```

## Chromagrams

### Euclidean chromagram

```{r}
chroma_plot_euclidian <- chroma_raw |>
  compmus_wrangle_chroma() |>
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
  compmus_gather_chroma() |>
  ggplot(aes(
    x = start + duration / 2,
    width = duration,
    y = pitch_class,
    fill = value
  )) +
  geom_tile() +
  labs(
    title = "Clocks (Coldplay) - Chromagram",
    x = "Time (s)", y = NULL, fill = "Magnitude"
  ) +
  theme_minimal() +
  scale_fill_viridis_c()

chroma_plot_euclidian
ggsave("plots/clocks_chromagram_euclidian.png", chroma_plot_euclidian, width = 10, height = 4, dpi = 300)
```

### Manhattan chromagram

```{r}
chroma_plot_manhattan <- chroma_raw |>
  compmus_wrangle_chroma() |>
  mutate(pitches = map(pitches, compmus_normalise, "manhattan")) |>
  compmus_gather_chroma() |>
  ggplot(aes(
    x = start + duration / 2,
    width = duration,
    y = pitch_class,
    fill = value
  )) +
  geom_tile() +
  labs(
    title = "Clocks (Coldplay) - Chromagram Manhattan",
    x = "Time (s)", y = NULL, fill = "Magnitude"
  ) +
  theme_minimal() +
  theme(panel.grid = element_blank()) +
  scale_fill_viridis_c() +
  scale_x_continuous(breaks = seq(0, 300, by = 20))

chroma_plot_manhattan
ggsave("plots/clocks_chromagram_manhattan.png", chroma_plot_manhattan, width = 10, height = 4, dpi = 300)
```

### Chebyshev chromagram

```{r}
chroma_plot_chebyshev <- chroma_raw |>
  compmus_wrangle_chroma() |>
  mutate(pitches = map(pitches, compmus_normalise, "chebyshev")) |>
  compmus_gather_chroma() |>
  ggplot(aes(
    x = start + duration / 2,
    width = duration,
    y = pitch_class,
    fill = value
  )) +
  geom_tile() +
  labs(
    title = "Clocks (Coldplay) - Chromagram Chebyshev",
    x = "Time (s)", y = NULL, fill = "Magnitude"
  ) +
  theme_minimal() +
  theme(panel.grid = element_blank()) +
  scale_fill_viridis_c() +
  scale_x_continuous(breaks = seq(0, 300, by = 20))

chroma_plot_chebyshev
ggsave("plots/clocks_chromagram_chebyshev.png", chroma_plot_chebyshev, width = 10, height = 4, dpi = 300)
```

## Cepstrogram

```{r}
cepstro_plot <- mfcc_raw |>
  compmus_wrangle_timbre() |> 
  mutate(timbre = map(timbre, compmus_normalise, "manhattan")) |>
  compmus_gather_timbre() |>
  ggplot(aes(x = start + duration / 2, width = duration, y = mfcc, fill = value)) +
  geom_tile() +
  labs(
    title = "Clocks (Coldplay) - Cepstrogram (Manhattan)",
    x = "Time (s)", y = NULL, fill = "Magnitude"
  ) +
  scale_fill_viridis_c() +
  theme_classic()

cepstro_plot
ggsave("plots/clocks_cepstrogram.png", cepstro_plot, width = 10, height = 4, dpi = 300)
```

## Self-similarity matrices

### Timbre self-similarity matrix

```{r}
timbre_ssm_plot <- mfcc_raw |>
  compmus_wrangle_timbre() |> 
  filter(row_number() %% 100L == 0L) |> 
  mutate(timbre = map(timbre, compmus_normalise, "manhattan")) |>
  compmus_self_similarity(timbre, "cosine") |> 
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width  = 100 * xduration,
      y = ystart + yduration / 2,
      height = 100 * yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(
    title = "Clocks (Coldplay) - Self-similarity (timbre)",
    x = "", y = ""
  )

timbre_ssm_plot
ggsave("plots/clocks_ssm_timbre.png", timbre_ssm_plot, width = 6, height = 6, dpi = 300)
```

### Chroma self-similarity matrix

```{r}
chroma_ssm_plot <- chroma_raw |>
  compmus_wrangle_chroma() |> 
  filter(row_number() %% 25L == 0L) |> 
  mutate(pitches = map(pitches, compmus_normalise, "manhattan")) |>
  compmus_self_similarity(pitches, "cosine") |> 
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = 25 * xduration,
      y = ystart + yduration / 2,
      height = 25 * yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(
    title = "Clocks (Coldplay) - Self-similarity (chroma)",
    x = "", y = ""
  )

chroma_ssm_plot
ggsave("plots/clocks_ssm_chroma.png", chroma_ssm_plot, width = 6, height = 6, dpi = 300)
```

## Interpretation

*Clocks* repeats the same piano pattern and chord loop for most of the song.  
You can see that in the chromagrams: the same few pitch classes keep showing up again and again across the track.  
The chroma self-similarity matrix also shows lots of similarity over time, which fits with the harmony staying mostly the same.  
The biggest exception is around **2:32**, where there is a clear break and the piano plays different notes or chords.

The timbre plots show section changes more clearly. The MFCC cepstrogram stays fairly consistent overall, but it shifts when the arrangement changes, such as when the drums first enter around **0:21**, when the vocals enter around **0:35**, when heavier drums come in around **1:20**, and during the break around **2:32**.

The timbre self-similarity matrix also shows clearer blocks than the chroma self-similarity matrix. This suggests that the overall sound and texture change between sections, even though the underlying chord loop often continues. In that sense, the chroma-based representations mainly highlight harmonic repetition, while the timbre-based representations are better at showing changes in instrumentation and arrangement.
