---
title: "Stop Drugs - Chromagram and Keygram"
format:
  html:
    theme: lumen
    toc: true
    toc-location: right
execute:
  echo: false
  warning: false
  message: false
---

## Introduction

In this section, I explore the pitch content of *Stop Drugs (In Me Gezicht) by Gotu Jim* using a chromagram, a keygram and a best-matching key over time plot. I choose this song, because in the last part of the track, something strange seems to happen to the pitch. I hope to find out more about that here.  

```{r}
library(tidyverse)
library(compmus)

stop_drugs <- read_csv("../songs_csv/stop_drugs_chroma.csv")
```

## Chromagram

The chromagram provides a time-based representation of pitch-class activity. It is useful for identifying recurring harmonic patterns and for getting a first impression of the tonal content of the song. 

```{r}
chromagram_plot <- stop_drugs |>
  compmus_wrangle_chroma() |>
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
  compmus_gather_chroma() |>
  ggplot(aes(
    x = start + duration / 2,
    width = duration,
    y = pitch_class,
    fill = value
  )) +
  geom_tile() +
  labs(
    title = "Stop Drugs (In Me Gezicht)- Chromagram",
    x = "Time (s)",
    y = NULL,
    fill = "Magnitude"
  ) +
  theme_minimal() +
  scale_fill_viridis_c()

chromagram_plot
```

As expected, we see a clear change in chroma features in the last part of the track, at around 150 seconds. This means that the distribution of prominent notes changes noticeably at that point. In musical terms, this suggests a shift in the harmonic content of the song, which may indicate a change of key or tonal center.

## Keygram

The keygram shows how well different major and minor key templates match the chroma features over time. To create the key templates, I first defined the pitch class distributions for a major and minor key based on the Krumhansl-Schmuckler key profiles. Then, I generated templates for all 12 major and 12 minor keys by circularly shifting these profiles.

```{r}
circshift <- function(v, n) {
  if (n == 0) v else c(tail(v, n), head(v, -n))
}

major_key <- c(6.35, 2.23, 3.48, 2.33, 4.38, 4.09, 2.52, 5.19, 2.39, 3.66, 2.29, 2.88)
minor_key <- c(6.33, 2.68, 3.52, 5.38, 2.60, 3.53, 2.54, 4.75, 3.98, 2.69, 3.34, 3.17)

key_templates <- tribble(
  ~name, ~template,
  "Gb:maj", circshift(major_key, 6),
  "Bb:min", circshift(minor_key, 10),
  "Db:maj", circshift(major_key, 1),
  "F:min", circshift(minor_key, 5),
  "Ab:maj", circshift(major_key, 8),
  "C:min", circshift(minor_key, 0),
  "Eb:maj", circshift(major_key, 3),
  "G:min", circshift(minor_key, 7),
  "Bb:maj", circshift(major_key, 10),
  "D:min", circshift(minor_key, 2),
  "F:maj", circshift(major_key, 5),
  "A:min", circshift(minor_key, 9),
  "C:maj", circshift(major_key, 0),
  "E:min", circshift(minor_key, 4),
  "G:maj", circshift(major_key, 7),
  "B:min", circshift(minor_key, 11),
  "D:maj", circshift(major_key, 2),
  "F#:min", circshift(minor_key, 6),
  "A:maj", circshift(major_key, 9),
  "C#:min", circshift(minor_key, 1),
  "E:maj", circshift(major_key, 4),
  "G#:min", circshift(minor_key, 8),
  "B:maj", circshift(major_key, 11),
  "D#:min", circshift(minor_key, 3))

```


```{r}
keygram_plot <- stop_drugs |>
  compmus_wrangle_chroma() |>
  filter(row_number() %% 25L == 0L) |>
  compmus_match_pitch_template(
    key_templates,
    method = "euclidean",
    norm = "euclidean"
  ) |>
  ggplot(aes(
    x = start + duration / 2,
    width = 50 * duration,
    y = name,
    fill = d
  )) +
  geom_tile() +
  scale_fill_viridis_c(guide = "none") +
  theme_minimal() +
  labs(
    title = "Stop Drugs - Keygram",
    x = "Time (s)",
    y = ""
  )

keygram_plot
```

The dark patches in the plot show keys with a relatively small euclidian distance to the chroma features at that time point, which indicates a better match. We can see that the keygram shows a clear change in the best-matching keys around 150 seconds, which corresponds to the change in chroma features observed in the chromagram. This suggests that there is indeed a shift in the tonal center of the song at that point.

## Best-matching key over time
To make the key estimation easier to interpret, I also extracted the single best-matching key at each time point instead of only plotting the full keygram. I first converted the chroma data into a format suitable for compmus and then reduced the time resolution by keeping every 50th frame, since key is generally more stable over longer stretches than individual chords.

Next, I compared each chroma vector to a set of major and minor key templates using compmus_match_pitch_template(). Because lower distance indicates a better match, I then selected the key with the smallest distance at each time point. Plotting these best-matching keys over time results in a simpler and more readable representation, which makes it easier to see whether the estimated key remains stable or changes during the track.

```{r}
stop_drugs |>
  compmus_wrangle_chroma() |>
  filter(row_number() %% 25L == 0L) |>
  compmus_match_pitch_template(
    key_templates,
    method = "euclidean",
    norm = "euclidean"
  ) |>
  group_by(start) |>
  slice_min(d, n = 1, with_ties = FALSE) |>
  ggplot(aes(x = start, y = name)) +
  geom_point() +
  geom_line(aes(group = 1)) +
  labs(
    title = "Best-matching key over time",
    x = "Time (s)",
    y = "Estimated key"
  ) +
  theme_minimal()


```

The plot of the best-matching key over time confirms the observations from the chromagram and keygram. There is a clear change in the estimated key around 150 seconds, which indicates a shift in the tonal center of the song. A cool thing to notice is the fact that this plot pretty much extracts the pattern of the blue patches from the keygram, but in a much more readable way. It also shows that the key estimation is not stable, as there are many fluctuations in the estimated key even before the change at 150 seconds. This is because the method is sensitive to small local changes in pitch-class content. If the harmony is somewhat ambiguous or not strongly centered on one key at every moment, several key templates can fit reasonably well. As a result, the best-matching key may fluctuate even when the overall tonal character of the song stays fairly similar.

## Conclusion

Overall, the three representations point in the same direction. The chromagram shows a clear change in pitch-class activity in the final part of the track, and both the keygram and the best-matching key plot suggest that this corresponds to a shift in tonal center around 150 seconds. At the same time, the key estimation remains somewhat unstable throughout, which shows that the song is not cleanly captured by a single key at every moment.

This also suggests a limitation of the keygram approach. In this case, the keygrams seem to be more useful for detecting changes in pitch or tonal center than for assigning one stable key with much confidence. For estimating key over a longer stretch of music, it may therefore be better to start from the chromagram itself, identify the most prominent pitch classes in that section, and then match those to the most likely key (F Major before 150 seconds and D Major afterwards). In that sense, the keygram is helpful as an exploratory tool, but the chromagram may be more reliable for interpreting the actual key over longer time intervals.